---
title: "BIOL 432 Assignment 1"
author: "Rishona Vemulapalli (20164619)"
date: "2023-01-13"
output: html_document
---

#### Link to my GitHub repository:
https://github.com/rishonav/BIOL432_Assignment1

- Final versions of R scripts (dataGenerato.R, volumEstimato.R, and ThemeScript.R) can be found in the scripts folder of the repository 
- Note: Some commit messages may be wrong as certain files got grouped together when committing/pushing to Github. 

### SETUP
```{r LoadingLibraries, message = FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
```

<br>

#### Running necessary R scripts sourced from computer

**Script #1:** Generates limb measurement data. First, vectors of each variable are created using `sample()` (Species, Observer; categorical data) and `rnorm()` (Limb length, Limb width; numeric data). All 4 vectors containing 100 randomly selected values each are combined into a single data frame. Using `write.csv`, the data frame is exported to a .csv file called measurements.csv. 
```{r GenerateDataScript}
source("./scripts/dataGenerato.R") 
```

<br>

**Script #2:** Calculates limb volume. Limb_width and Limb_length measurements from measurements.csv are used to calculate limb volume (visualized as a cone shape). The Volume variable is added to the data frame. Using `write.csv` again, the measurements.csv is overwritten with the updated dataframe.  
```{r VolEstimatorScript}
source("./scripts/volumeEstimato.R")
```

<br>

**Script #3:** Clean theme to use for `ggplot` plotting (originally sourced from http://bit.ly/theme_pub). 
```{r ThemeScript}
source("./scripts/ThemeScript.R") 
theme_set(theme_pub()) # setting the source theme as default
```

<br>

### QUESTIONS 

#### (1) Loading the overwritten data

```{r Import}
MeasurementData <- read.csv("measurements.csv")
print(MeasurementData)
```

<br>

#### (2) Sorting data by species, then observer, and then limb volume

```{r Reorder}
MeasurementData <- MeasurementData %>%
  arrange(Species, Observer, Volume) %>%
  print()
```

<br>

#### (3) Table of average limb volume by species

```{r LimbSummary}
LimbSummary <- MeasurementData %>%
  group_by(Species) %>%
  summarise("Average Volume" = mean(Volume)) %>%
  print()
```

<br>

#### (4) Table of number of observations (rows) for each combination of species and observer

```{r CountSummary}
CountSummary <- MeasurementData %>%
  count(Species, Observer) %>%
  rename(Num_Observations = n) %>% # number of observations
  print()
```

<br>

#### (5) Single boxplot to compare distributions of limb volume for each species

```{r Fig1}
Limb_bp <- ggplot(MeasurementData, aes(x = Species, y = Volume)) +
  geom_boxplot(aes(fill = Species)) +
  scale_x_discrete(labels = c("D. versicolor", "L. clamitans", "L. pipiens", "P. crucifer", "P. maculata")) +
  scale_y_continuous("Limb volume (cm^3)", limits = c(0, 75), breaks = c(0, 25, 50, 75)) +
  scale_fill_brewer(palette = "Set2")
print(Limb_bp)
```

Figure 1. Boxplot of limb volume (cm^3) of 5 frog species; *Dryophytes versicolor*, *Lithobates clamitans*, *Lithobates pipiens*, *Pseudacris crucifer*, *Pseudacris maculata*). Limb volume was calculated using measurements of limb width and limb length. 

<br>

#### (6) Multi-panel frequency histograms plot for limb volume of each species

Creating the multi-panel plot using `facet_wrap`. 
```{r Fig2, message = FALSE}
Limb_hists <- ggplot(MeasurementData, aes(x = Volume)) +
  geom_histogram(aes(fill = Species), colour = "black") +
  facet_wrap(Species ~ .) +
  labs(x = "Limb volume (cm^3)", y = "Count") +
  scale_fill_brewer(palette = "Set2")
print(Limb_hists)
```

Figure 2. Multi-panel plot of 5 frequency histograms, showing the distribution of limb volume (cm^3) for each of the 5 frog species. Fixed x- and y-scales to compare distributions between levels.   
