---
title: "BIOL 432 Week 5 A4: R-egex"
author: "Rishona Vemulapalli (20164619)"
date: "2023-02-08"
output: html_document
---

### Link to my GitHub Repository:
https://github.com/rishonav/BIOL432_Assignment1

-------------------------------------------------------------

### (SETUP)

```{r message = FALSE}
library(tidyverse)
```

-------------------------------------------------------------

### (QUESTION 1) Import dataset created from Assignment 1.

*Note: The Organism column named in this assignment's instructions is the Species column in my dataset from Assignment 1.*
```{r}
MData <- read.csv("measurements.csv")
head(MData)
```

<br>

-------------------------------------------------------------

### (QUESTION 2) Add a new column called *sp* that shortens genus name in Species column to only first letter then period. 

- **Use a dplyr command AND regular expression(s)**
- **For example: Homo sapiens original column becomes H. sapiens in Sp column.**
- **HINT: Think of regular expression commands like sub, grep, grepl and gsub as 'functions' that you can use with dplyr commands just like mean(), sd(), and desc() in
the Data Science tutorial.**

In a new added column called Sp, I am replacing the first match of the Genus name, comprised of any alphanumeric characters, in the *Species* column, with its first letter followed by a period, and its species name. 
```{r}
MDataShorter <- MData %>%
  mutate(Sp = sub("(\\w)\\w+", "\\1.", Species), .after = Species) # placing new Sp column after the Species (Organism) column to make selection easier after

print(MDataShorter) 
```

<br>

-------------------------------------------------------------

### (QUESTION 3) Create a new dataset with the Sp column. 

- **Use a dplyr command to create a new dataset BUT with Sp column instead of the Organism column, moved to the beginning in place of the original Organism column.**

```{r}
MeasShortData <- select(MDataShorter, Sp:Volume)
print(MeasShortData)
```

<br>

-------------------------------------------------------------

### (QUESTION 4) Save the dataset as a new file called MeasShort.csv

```{r}
write.csv(MeasShortData, file = "MeasShort.csv")
```

<br>
